Publilia_2019_CHAPTER-0-prep

### 2023-08-24 scripts/old/Publilia_2019_CHAPTER-0-prep.Rmd
added old scripts to `scripts/old/` for reference

---
title: '01 - TreehopperFitness: Quality control'
author: "Micah Fletcher"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
  fig_width: 9
toc_depth: 3
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 6)

library(kableExtra)   ## Needed if rendering tables
```

```{=html}
<style type="text/css">
  .book .book-body .page-inner {
    max-width: 1600px;
    margin-left: auto;
    margin-right: auto;
  }
details > summary {
  display: list-item;
}
</style>
```

# Data input, qc and pre-processessing for 2019 Publilia field data
```{r}
library(here)
library(tidyverse)
library(readxl)
library(lubridate)
library(stringr)

theme_set(theme_bw())
```

## 0 - Raw Data Input
Read in the data from `data_in_progress/Watershed_field_data_2019_ALL_local.xlsx`
```{r, warning = FALSE}
col_types = c("date", rep("text",8), "date", rep("text", 12))
in_filename <- here("data/Watershed_field_data_2019_ALL_local.xlsx")
in_data <- rbind(read_xlsx(in_filename, sheet = "A", col_types = col_types),
  read_xlsx(in_filename, sheet = "B", col_types = col_types),
  read_xlsx(in_filename, sheet = "C", col_types = col_types),
  read_xlsx(in_filename, sheet = "D", col_types = col_types),
  read_xlsx(in_filename, sheet = "E", col_types = col_types),
  read_xlsx(in_filename, sheet = "F", col_types = col_types),
  read_xlsx(in_filename, sheet = "G", col_types = col_types),
  read_xlsx(in_filename, sheet = "H", col_types = col_types)) %>%
  mutate(Date = as_date(Date),
         Time = hm(Time)) %>% 
  group_by(row_number()) %>% 
  mutate(Leaf = ifelse(!is.na(Female_ID) & is.na(Leaf), 
                       "S",
                       Leaf),
         Female_ID = ifelse(Female_ID %in% c("Uf", "Um", "M", "U") , # append plant and leaf to ID if female id is "unknown"
                            paste0(c(Female_ID, "_", as.character(Plant), "_", as.character(Leaf)), collapse = ""),
                            Female_ID)) %>%
  ungroup()
```

# 1 - Manually check and revise female IDs reproducibly

Create `Female_locations_on_dates` table to use for to manually check for obvious mismatched female IDs and typos in plant id or leaf id etc.

Then manually log these changes into `{dorm letter} Female_ID revisions` sheet then import these and revise the female IDs in the input data without editing the original data file.

Note that changes to things like plant id and leaf id had to be manually edited in the original data file. These are documented in the "Revisions" column in each sheet for each dorm in `Watershed_field_data_2019_ALL_local.xlsx`

## 1.1 - Flag duplicate female ID entries

Flag cases where the same female ID is used on the same day in the same dorm to aid in manual checking.

Display false negative and false positive duplicates in `Dup?` column (used in excel for field data entry).

```{r}
in_data_dupStatus <- in_data %>%
  group_by(Date, Dorm, Female_ID) %>%
    mutate(n_female_dupes = ifelse(!is.na(Female_ID) & !(Female_ID %in% c("E", "E?", "0", NA)),
                                   n(),
                                   NA),
           female_dup_order = ifelse(!is.na(Female_ID) & !(Female_ID %in% c("E", "E?", "0", NA)),
                                     row_number(),
                                     NA)) %>%
  ungroup() %>%
  mutate(dup_status = ifelse(!is.na(n_female_dupes) & !(Female_ID %in% c("E", "E?", "0", "N", NA)),
                             ifelse(n_female_dupes == 1 & !is.na(as.numeric(`Dup?`)),
                                    "Fpos", # copy-paste error or "?" ambiguity in Excel
                                    ifelse(n_female_dupes > 1 & !is.na(as.numeric(`Dup?`)),
                                           "Tpos", 
                                           ifelse(n_female_dupes > 1 & is.na(as.numeric(`Dup?`)),
                                                  "Fneg", # copy-paste error
                                                  ifelse(n_female_dupes == 1 & is.na(as.numeric(`Dup?`)),
                                                         "Tneg",
                                                         "Something went wrong!")))),
                             NA),
         Fem_IDn = ifelse(!is.na(n_female_dupes) & !(Female_ID %in% c("E", "E?", "0", "N", NA)),
                               ifelse(n_female_dupes == 1,
                                      Female_ID,
                                      paste0(Female_ID, "_d", female_dup_order)),
                               NA))

in_data_dupStatus %>%
  count(dup_status)

@@

Days_observed_dupStatus <- female_roster_days_dupStatus %>% 
  group_by(Dorm, Fem_IDn) %>%
    summarize(Days_observed = n()) %>%
  ungroup() %>%
  arrange(Dorm, Days_observed)

# used 20200319 version
# write_csv(Days_observed_dupStatus, here("data/manual_check/Days_observed_dupStatus.csv"))
```

## 1.3 - Revise Female_ID using `Female_ID revisions dupStatus` sheet

```{r}
# Read in revisions table
Female_ID_revisions <- read_xlsx(in_filename, sheet = "Female_ID revisions dupStatus", col_types = c("date", "text", "text", "text","text", "text", "text", "text"), na = c("", "NA")) %>%
  mutate(Date = as_date(Date),
         Female_ID_revision_notes = Notes) %>%
  select(-Notes) %>%
  mutate(Date = `year<-`(Date, 2019)) # excel's fuckery with dates means any revisions I did in 2020 in the format m/d were interpretted as 2020-%m-%d

@@

*Note: This section saves many `.csv` files with tables to aid manual checks. In each case, the contents of these files were copied into the `data/Watershed_field_data_2019_ALL_local.xlsx` spreadsheet as new sheet at the time of initial analysis, revised manually as needed and subsequently read back in to this analysis script and used to revise the data object in the R environment. Every effort was made to keep the original field-collected dataseets unchanged in favor of programmatically revising the loading R object. In a few cases where this was not possible, the original spreadsheet was edited (with a note in the `Revisions` column). As such, newly generated check tables using this script may differ from the originals documented in `data/Watershed_field_data_2019_ALL_local.xlsx`, which preserve the original state of the data.*

@@

## 2.2 - Create tables to use for manually checking and revising egg status text conflict or missing data
```{r}
# Create table for entries text conflict that need to be resolved
egg_text_conflict_data_v1 <- All_data_v2 %>% # egg_text_conflict: 39 on 2019/09/24
  left_join(egg_status_v1, by = c("Date", "location", "Eggs" = "Eggs_status")) %>%
  filter(egg_text_conflict) %>%
  select(Date, Dorm, Plant, Leaf, Eggs)
# write_csv(egg_text_conflict_data_v1, here("data/manual_check/egg_text_conflict_data.csv"))

# Create table for entries with missing data that need to be resolved
egg_status_missing_data_v1 <- All_data_v2 %>% # egg_status_missing: 219 on 2019/09/24
  left_join(egg_status_v1, by = c("Date", "location", "Eggs" = "Eggs_status")) %>%
  filter(egg_status_missing, !grepl('NA|M', Female_ID)) %>%
  select(Date, Dorm, Plant, Leaf, Eggs)
# write_csv(egg_status_missing_data_v1, here("data/manual_check/egg_status_missing_data.csv"))

@@

# 20200326 version was used
# write_csv(egg_status_missing_data_eggs_revised, here("data/manual_check/egg_status_missing_data_eggs_revised.csv"))
```

@@

# 20200326 version was used
# write_csv(egg_status_missing_data_eggs_revised, here("data/manual_check/egg_status_missing_data_eggs_revised.csv")
```

@@

nymph_status_to_resolve <- nymph_status  %>% # should only include a handful of entries with "?" ambiguous entries for N1 or N2 (because did the revions by hand on 2019/09/26)
  group_by(row_number()) %>%
    filter(any(N1_need_filling, N1_text_conflict, N1_status_missing,
               N2_need_filling, N2_text_conflict, N2_status_missing,
               N5_need_filling, N5_text_conflict, N5_status_missing)) %>%
  ungroup()
# write_csv(nymph_status_to_resolve, here("data/manual_check/nymph_status_to_resolve.csv") ## manually added this to "nymph_status revisions" sheet in "Watershed_field_data_2019_ALL_local" and manually revised raw data 2019/09/26

@@

# checking number of ambiguous entries that need to be resolved
nymph_text_conflict_data <- All_data_v4 %>% # nymph_text_conflict: 10 on 2019/09/24; 0 on 2019/09/26
  left_join(N1_status, by = c("Date", "location", "N1")) %>%
  left_join(N2_status, by = c("Date", "location", "N2")) %>%
  left_join(N5_status, by = c("Date", "location", "N5")) %>%
  group_by(row_number()) %>% 
    filter(any(c(N1_text_conflict, N2_text_conflict, N5_text_conflict))) %>%
  ungroup() %>%
  select(Date, Dorm, Plant, Leaf, Female_ID, Eggs, N1, N2, N5, nymph_origin, Notes, N1_text_conflict, N2_text_conflict, N5_text_conflict)
# write_csv(egg_text_conflict_data, here("data/manual_check/egg_text_conflict_data.csv"))

nymph_status_missing_data <- All_data_v4 %>% # nymph_status_missing: 17 on 2019/09/24; 5 on 2019/09/26 (all ok to ignore)
  left_join(N1_status, by = c("Date", "location", "N1")) %>%
  left_join(N2_status, by = c("Date", "location", "N2")) %>%
  left_join(N5_status, by = c("Date", "location", "N5")) %>%
  group_by(row_number()) %>% 
  filter(any(c(N1_status_missing, N2_status_missing, N5_status_missing))) %>%
  ungroup() %>%
  select(Date, Dorm, Plant, Leaf, Female_ID, Eggs, N1, N2, N5, nymph_origin, Notes, N1_status_missing, N2_status_missing, N5_status_missing)
# write_csv(egg_status_missing_data, here("data/manual_check/egg_status_missing_data.csv")

@@

## 2.9 - Manually check hatch dates using the "not hatched" notes from field using `check_hatch_date` sheet
Use the "not hatched" notes (from original field notes; typically b/c nymphs present but they are immigrants, not hatched from the focal clutch) to inform first pass of hatch date revisions
Again these changes were specified in the `check_hatch_date` sheet and are applied here without changing the original data file.
```{r}
# Create `check_hatch_date` sheet for checking "not hatched" notes and suggesting manual revisions
check_hatch_date <- All_data_v4  %>%
  filter(grepl("not", Notes, ignore.case = TRUE), grepl("hatch", Notes, ignore.case = TRUE)) %>%
  distinct(location)
# write_csv(check_hatch_date, here("data/manual_check/check_hatch_date_tmp.csv")
```

@@

# write_csv(ant_text_conflict, here("data/manual_check/ant_text_conflict.csv") # resolved 2019-10-08
```

@@

# write_csv(short_unhatched_days_check, here("data/manual_check/short_unhatched_days_check_tmp.csv")

@@

# write_csv(never_hatched_check, here("data/manual_check/never_hatched_check_tmp.csv")

# write_csv(clutch_status_stats_v1, here("data/manual_check/egg_measures_check_tmp.csv")
```

@@

***

*Prep data for `Publilia-03-survivorship.Rmd`:*

# 7 - Assigning nymph origins

Nymphs migrate away from their natal leaf. On plants with multiple clutches, it was sometimes possible to determine the origin of nymphs later in the season because sometimes groups of nymph don't travel very far from their natal leaf and mix with other groups. I manually curated examples of this, so that clutch-level stats on hatching success and nymph survivorship could be determined on the plants with multiple clutches. Some clutches (as a rule only on multi-clutch plants) did not have a clear record of nymphs that emigrated from the natal leaf and thus could not be included in future clutch-level analyses of survivorship.

## 7.1 - Adding second gen adult counts per leaf to clutch status stats
```{r}
# counting adults
eA_status_v1 <- All_data_v6 %>%
  group_by(row_number()) %>% # sum of newly exclosed adults
  filter(any(!is.na(c(New_eA, Old_eA)))) %>%
  mutate(n_newA = ifelse(is.na(New_eA),
                         NA,
                         sum(as.numeric(regmatches(New_eA, gregexpr("[0-9]+", New_eA))[[1]]))),
         n_oldA = ifelse(is.na(Old_eA),
                         NA,
                         sum(as.numeric(regmatches(Old_eA, gregexpr("[0-9]+", Old_eA))[[1]]))),
         n_A = sum(n_newA, n_oldA, na.rm = TRUE)) %>%
  ungroup() %>%
  select(Date, location, n_newA, n_oldA, n_A)

All_data_v6_clutch_status_stats_v2 <- All_data_v6 %>%
  # add egg measures, clutch status stats (including initiation dates, hatch dates, egg fate, etc)
  left_join(clutch_status_v2, by = c("location", "Date")) %>%
  left_join(clutch_status_stats_v1, by = c("location")) %>%
  left_join(eA_status_v1, by = c("Date", "location")) %>%
  left_join(select(clutch_status_guarding_summary_v2, location, never_hatched)) %>%
  mutate(N1 = as.numeric(N1),
         N2 = as.numeric(N2),
         N5 = as.numeric(N5))
```

## 7.2 - Manually check nymph origins for revision
First create a table of all entried from clutches that were not flagged in the field as having ambiguous origin (e.g. "leaf 6 or 9") in cases were more than one clutch was on the plant. These were revised manually to specify a leaf of origin for the nymphs, where possible.
```{r}
NOr_status_resolveable_all_unambig_clutches <- All_data_v6_clutch_status_stats_v2 %>% 
  select(Date, Dorm, Plant, Leaf, location, 
         egg_num, N1, N2, N5, n_A, nymph_origin, initiation_date, hatch_date,
         any_NOr_ambig_loc, Omit_for_abandon_time, never_hatched,
         n_clutches_on_plant) %>%
  # keep only entries with some nymphs or adults on a leaf
  # OR entries for leaves where clutches have been initiated
  group_by(row_number()) %>%
  filter(!is.na(Leaf) & any(N1 > 0, N2 > 0, N5 > 0, n_A > 0) | 
           Date >= initiation_date) %>% # nrow = 17944
  # flag entries where the nymph origin is text (typically ambiguous, e.g. "8 or 9")
  mutate(NOr_ambig_entry = !is.na(nymph_origin) & is.na(as.numeric(nymph_origin))) %>% 
  ungroup() %>%
  # get rid of duplicate rows from multiple females in the same location
  distinct(Date, location, .keep_all = TRUE) %>% # nrow = 15674
  group_by(Dorm, Plant) %>%
    # filter out any plants with a clutch that was omitted for abandon time
    filter(!any(!is.na(Omit_for_abandon_time))) %>%
    # flag plants with any ambiguous entries (e.g. "8 or 9") skip for checking
    mutate(NOr_ambig_plant = any(NOr_ambig_entry)) %>%
  ungroup()

NOr_status_resolveable_mult_clutches <- NOr_status_resolveable_all_unambig_clutches %>% # nrow = 15674
  group_by(Dorm, Plant) %>%
    # plants with mutliple clutches only
    filter(any(n_clutches_on_plant > 1)) %>%
  ungroup() %>%
  group_by(row_number()) %>%
    # flag for manual checking entries with any nymphs/adults, missing nymph origin, 
    # but not flagged for ambiguous origin in the field.
    # these are the most likely to be easily resolved with minimal work
    mutate(revise_NOr = ifelse(any(N1 > 0, N2 > 0, N5 > 0, n_A > 0) &
             !NOr_ambig_plant,
           "MANUAL CHECK",
           NA)) %>% 
  ungroup()

NOr_manual_check_table <- NOr_status_resolveable_mult_clutches %>%
  group_by(Dorm, Plant) %>%
    filter(any(revise_NOr == "MANUAL CHECK")) %>% # nrow = 2830
  ungroup() %>%
    group_by(row_number()) %>%
    filter(any(N1 > 0, N2 > 0, N5 > 0, n_A > 0)) %>%
  ungroup() %>%
  select(Date, Dorm, Plant, Leaf, location, egg_num, N1, N2, N5, n_A, nymph_origin,
         initiation_date, hatch_date,
         revise_NOr) %>%
  arrange(Dorm, as.numeric(Plant), as.numeric(Leaf), Date)
# 83 clutches on 34 plants to check

# write_csv(NOr_manual_check_table %>% filter(revise_NOr == "MANUAL CHECK"), here("data/manual_check/NOr_manual_check_tmp.csv")

NOr_manual_revisions <- read_xlsx(in_filename, sheet = "NOr_manual_check", col_types = c("date", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "text", "logical", "text")) %>%
  mutate(Date = as_date(Date),
         NOr_rev_loc_manual = paste0(Dorm, Plant, "_", NOr_revised)) %>%
  filter(!cant_resolve) %>%
  select(Date, location, NOr_rev_loc_manual, NOr_revised_manual = NOr_revised)

NOr_manual_revisions %>%
  distinct(NOr_rev_loc_manual) # 43/83 initiated clutches (43/59 hatched) with successfully revised NOr
```

## 7.3 - Implement revisions by creating the new variables NOr_revised and NOr_rev_loc
```{r}
# if nymph origin was clearly resolveable, the revisions were made in NOr_manual_revisions
# updating NOr_status with revisions so that it can be left_joined to other data and
# counts of nymphs and adults on NOr can be calculated
NOr_status_revised_all_unambig_clutches <- NOr_status_resolveable_all_unambig_clutches %>%
  left_join(NOr_manual_revisions, by = c("Date", "location")) %>%
  group_by(Dorm, Plant) %>%
    mutate(lone_hatcher_leaf = ifelse(length(unique(Leaf[!is.na(initiation_date)])) == 1,
                                 unique(Leaf[!is.na(initiation_date)]),
                                 NA),
           NOr_revised = ifelse(!is.na(NOr_revised_manual),
                                       NOr_revised_manual,
                                ifelse(any(!is.na(lone_hatcher_leaf)),
                                       lone_hatcher_leaf,
                                       # if no clutches hatched on the plant, NOr_revised = Leaf
                                       # for leaves with eggs
                                       ifelse(!any(!is.na(hatch_date)) & !is.na(egg_num),
                                              Leaf,
                                              NA)))) %>%
  ungroup() %>%  
  group_by(location) %>%
    mutate(NOr_revised = ifelse(any(NOr_revised == Leaf) & is.na(NOr_revised),
                                Leaf,#unique(NOr_revised[!is.na(NOr_revised)]),
                                NOr_revised)) %>%
  ungroup() %>%
  mutate(NOr_rev_loc = ifelse(is.na(NOr_revised),
                              NA,
                              paste0(Dorm, Plant, "_", NOr_revised))) %>%
  select(Date, location, NOr_revised, NOr_rev_loc) %>%
  filter(!is.na(NOr_rev_loc))

All_data_v6_N_NOr_counts_unambig_clutches <- All_data_v6_clutch_status_stats_v2 %>%
  right_join(NOr_status_revised_all_unambig_clutches, by = c("Date", "location")) %>%
  group_by(Date, Dorm, Plant, NOr_revised, NOr_rev_loc) %>%
    summarize(N1_NOr = sum(N1, na.rm= TRUE),
              N2_NOr = sum(N2, na.rm= TRUE),
              N5_NOr = sum(N5, na.rm= TRUE),
              N_NOr = sum(N1, N2, N5, na.rm = TRUE),
              A_NOr = sum(n_A, na.rm= TRUE),
              N5plusA_NOr = sum(N5, n_A, na.rm= TRUE)) %>%
  ungroup() # n = 186

All_data_v6_N_NOr_counts_unambig_clutches_summary <- All_data_v6_N_NOr_counts_unambig_clutches %>%
  group_by(Date) %>%
    mutate(mean_N_NOr = mean(N_NOr, na.rm = TRUE),
           mean_N1_NOr = mean(N1_NOr, na.rm = TRUE),
           mean_N2_NOr = mean(N2_NOr, na.rm = TRUE),
           mean_N5_NOr = mean(N5_NOr, na.rm = TRUE),
           mean_A_NOr = mean(A_NOr, na.rm = TRUE),
           mean_N5plusA_NOr = mean(N5plusA_NOr, na.rm = TRUE),
           All_N_NOr = sum(N_NOr),
           All_A_NOr = sum(A_NOr)) %>%
  ungroup() %>%
  group_by(Date, Dorm) %>%
    mutate(All_N_NOr_Dorm = sum(N_NOr),
           All_N1_NOr_Dorm = sum(N1_NOr),
           All_N2_NOr_Dorm = sum(N2_NOr),
           All_N5_NOr_Dorm = sum(N5_NOr),
           All_A_NOr_Dorm = sum(sum(A_NOr))) %>%
  ungroup()

All_data_v6_all_nymphs_unambig_clutches <- All_data_v6_N_NOr_counts_unambig_clutches %>%
  group_by(Date) %>%
    summarize(All_N_NOr = sum(N_NOr)) %>%
  ungroup()

All_data_v6_max_N_NOr_counts_unambig_clutches <-
  All_data_v6_N_NOr_counts_unambig_clutches %>%
  group_by(Dorm, Plant, NOr_revised, NOr_rev_loc) %>%
    summarize(max_N_NOr = max(N_NOr)) %>%
  ungroup()

>>>>

---
title: "Publilia_2019_CHAPTER-0-prep"
author: "micah fletcher"
date: "1/18/2022"
output: html_document
---

# Publilia 2019 Chapter - Script 0: data prep
This is meant to consolidate previous scripts and provide a readable and reproducible record of the analyses for the Publilia Chapter.

The outline for the chapter and the corresponding scripts are as follows:

0) *Data pre-processing:* 
Preprocessing necessary for all the following analyses, including various automated and manual checking procedures (Female IDs, numbers of eggs, nymphs, ants etc.).

1) *Guarding strategies:* Three distinct strategies for maternal care emerge from ant-dependent and secondary visitor-dependent plasticity
  - *Dataset A* requirements:
    - Guarding durations (relative to [estimated] hatching) are accurate.
      - The sample of clutches is representative of the full factorial of ants / no ants and visited early / not visited early
    - Able to compare 
      - initiated vs. visited
      - first vs. later [initiated] clutches
        -however, if plotting "all clutches" I can include unmarked females
      - disappeared vs. deserted (based on last day a female was observed)
    - Deciding on who is the "true" initiator is necessary for assumptions of independence for statistics. 
      - Ultimately, comparing the two most obvious options (first and longest guarding mom at initiation, vs first and shortest guarding mom at initiation) will need to suffice.
Missing data is clearly represented; Compare the "original" unfiltered data to the final data and report why (and how many) data were excluded.

2) *Fecundity:* No support for trade-off between current care and future fecundity
  - *Dataset B [supp: modeling fecundity ~ guarding]* requirements:
    - Guarding durations of ALL females that visited a clutch are accurate.
      - Cannot exclude any females, including unmarked or ambiguously marked females, because I need them for the sum of guarding time.
    - Maximum number of eggs per clutch is accurate.
    - Can partition 
      - early clutches vs. later clutches
      - single-mother clutches vs. multiple-mother clutches
      - ants vs. no ants
  - *Dataset C [estimating fecundity]* requirements:
    - Guarding durations of ALL females that visited a clutch are accurate.
      - Cannot exclude any females, including unmarked or ambiguously marked females, because I need them for the sum of guarding time.
    - Maximum number of eggs per clutch is accurate.
    - Numbers of clutches per female (and guarding duration and eggs for each) are accurate.
      - Cannot exclude any clutches, including those with ambiguous initiators, or unknown hatch dates
    - Models/heuristics of focal female contribution to clutch are clearly defined and sensible.
      1) simple equal proportion of days
      2) equal proportion of days with "intercept" given to the initiator
      3) unequal proportion of days for initiator(s) vs visitors, with intercept also given to the initiator
      4) Reproducing Zink's fecundity

3) *Clutch-level survivorship regressions:* The impact of care on offspring survivorship depends on ants and offspring stage
  - *Dataset D [survivorship ~ guarding duration]* requirements:
    - Guarding durations (relative to [estimated] hatching) are accurate.
    - (Clutch-level) Hatching success is accurate.
    - Number of eggs is accurate.
    - Must exclude clutches where migration to/from other leaves obscured nymph count.
      - This biases the results toward only including clutches alone on a plant, while clearly it is common to have multiple clutches on a plant. One way to address this is to compare to plant-level analysis, so I may reuse parts of this pipeline for the plant-level analysis.
    - (Clutch-level) Nymph survivorship is accurate.
    - Must exclude clutches where migration to/from other leaves obscured nymph count. This biases the results in the same way as above.
    - Main analysis on first clutches laid before June 10. 
    - But also report on [] in the supplement.
      - the full dataset
      - single-female clutches only
      - female-removal included vs. excluded
    - Missing data is clearly represented; Compare the "original" unfiltered data to the final data and report why (and how many) data were excluded.
  - *Dataset E [clutch survivorship ~ ants]* requirements:
    - Same as Dataset B but also need a representative measure of ant attendance intensity/persistence, e.g.:
      - percent days attended before hatch
      - mean/median number of ants before hatch
      - number of days it took for ants to appear at clutch after hatch
      - Note: these would ideally be comparable to plant-level metrics

4) *Plant-level survivorship:* Do nymph migration and ant availability at a plant-level make plant-level dynamics more informative to survivorship than clutch-level?
  - Dataset F [survivorship ~ female presence + ant presence + number of clutches on plant]
most of the same metrics as for datasets D and E but with relaxed filtering of clutches with ambiguous nymph origins

5) *Ovary development:* If I have time (Narrator: He won't)

## Notes on previous versions.
Started quality checking in Fall 2019 in `female_ID_qc.R`. Started rewriting code in Rmd format on 3/11/2020 in `Watershed_2019_Analysis_IP.Rmd`. Started reformatting code into multiple scripts, separating data prep from analysis for better readability, reproducibility, and modularity on 4/13/2021. This final version was started 1/18/2022.

# Data input, qc and pre-processessing for 2019 Publilia field data
```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(stringr)

theme_set(theme_bw())

setwd("/Users/micah/Documents/Princeton/Kocher_Lab/Publilia/PubliliaBehavEcol/Watershed_2019/Watershed_2019_analysis/CHAPTER")
```

## 0 - Raw Data Input
Read in the data from `data_in_progress/Watershed_field_data_2019_ALL_local.xlsx`
```{r, warning = FALSE}
col_types = c("date", rep("text",8), "date", rep("text", 12))
in_filename <- "../data_in_progress/Watershed_field_data_2019_ALL_local.xlsx"
in_data <- rbind(read_xlsx(in_filename, sheet = "A", col_types = col_types),
  read_xlsx(in_filename, sheet = "B", col_types = col_types),
  read_xlsx(in_filename, sheet = "C", col_types = col_types),
  read_xlsx(in_filename, sheet = "D", col_types = col_types),
  read_xlsx(in_filename, sheet = "E", col_types = col_types),
  read_xlsx(in_filename, sheet = "F", col_types = col_types),
  read_xlsx(in_filename, sheet = "G", col_types = col_types),
  read_xlsx(in_filename, sheet = "H", col_types = col_types)) %>%
  mutate(Date = as_date(Date),
         Time = hm(Time)) %>% 
  group_by(row_number()) %>% 
  mutate(Leaf = ifelse(!is.na(Female_ID) & is.na(Leaf), 
                       "S",
                       Leaf),
         Female_ID = ifelse(Female_ID %in% c("Uf", "Um", "M", "U") , # append plant and leaf to ID if female id is "unknown"
                            paste0(c(Female_ID, "_", as.character(Plant), "_", as.character(Leaf)), collapse = ""),
                            Female_ID)) %>%
  ungroup()

# 1 - Manually check and revise female IDs reproducibly
Create `Female_locations_on_dates` table to use for to manually check for obvious mismatched female IDs and typos in plant id or leaf id etc.
Then manually log these changes into `{dorm letter} Female_ID revisions` sheet then import these and revise the female IDs in the input data without editing the original data file.
Note that changes to things like plant id and leaf id had to be manually edited in the original data file. These are documented in the "Revisions" column in each sheet for each dorm in `Watershed_field_data_2019_ALL_local.xlsx`

## 1.1 - Flag duplicate female ID entries
Flag cases where the same female ID is used on the same day in the same dorm to aid in manual checking.
Display false negative and false positive duplicates in `Dup?` column (used in excel for field data entry).
```{r}
in_data_dupStatus <- in_data %>%
  group_by(Date, Dorm, Female_ID) %>%
    mutate(n_female_dupes = ifelse(!is.na(Female_ID) & !(Female_ID %in% c("E", "E?", "0", NA)),
                                   n(),
                                   NA),
           female_dup_order = ifelse(!is.na(Female_ID) & !(Female_ID %in% c("E", "E?", "0", NA)),
                                     row_number(),
                                     NA)) %>%
  ungroup() %>%
  mutate(dup_status = ifelse(!is.na(n_female_dupes) & !(Female_ID %in% c("E", "E?", "0", "N", NA)),
                             ifelse(n_female_dupes == 1 & !is.na(as.numeric(`Dup?`)),
                                    "Fpos", # copy-paste error or "?" ambiguity in Excel
                                    ifelse(n_female_dupes > 1 & !is.na(as.numeric(`Dup?`)),
                                           "Tpos", 
                                           ifelse(n_female_dupes > 1 & is.na(as.numeric(`Dup?`)),
                                                  "Fneg", # copy-paste error
                                                  ifelse(n_female_dupes == 1 & is.na(as.numeric(`Dup?`)),
                                                         "Tneg",
                                                         "Something went wrong!")))),
                             NA),
         Fem_IDn = ifelse(!is.na(n_female_dupes) & !(Female_ID %in% c("E", "E?", "0", "N", NA)),
                               ifelse(n_female_dupes == 1,
                                      Female_ID,
                                      paste0(Female_ID, "_d", female_dup_order)),
                               NA))

@@

Days_observed_dupStatus <- female_roster_days_dupStatus %>% 
  group_by(Dorm, Fem_IDn) %>%
    summarize(Days_observed = n()) %>%
  ungroup() %>%
  arrange(Dorm, Days_observed)
#write_csv(Days_observed_dupStatus, "Days_observed_dupStatus_20200319.csv")
```

## 1.3 - Revise Female_ID using `Female_ID revisions dupStatus` sheet
- ~~check that each "From ID" is actually in the same location in the data as on the revisions table, flag where not the case~~ *[DONE]*
- ~~check that all duplicates have been resolved~~ *[DONE]*
- ~~manually doublecheck the location on date table (especially the entry for E_Rgr?_d1)~~ *[DONE]*
- *****manually check that `short_unhatched_days_check` female revisions remain appropriate and make these revisions reversible -> Check this after running the "short_unhatched_days_check" code chunk*****
-  ~~old method of replacing a NA (in `From ID`) with a Female_ID (in `To ID`) doesn't work; I will make a new `NA to ID` column that can add a Female ID where there was an NA before~~ *[DONE]*
```{r}
# Read in revisions table
Female_ID_revisions <- read_xlsx(in_filename, sheet = "Female_ID revisions dupStatus", col_types = c("date", "text", "text", "text","text", "text", "text", "text"), na = c("", "NA")) %>%
  mutate(Date = as_date(Date),
         Female_ID_revision_notes = Notes) %>%
  select(-Notes) %>%
  mutate(Date = `year<-`(Date, 2019)) # excel's fuckery with dates means any revisions I did in 2020 in the format m/d were interpretted as 2020-%m-%d

@@

@@

## 2.2 - Create tables to use for manually checking and revising egg status text conflict or missing data
```{r}
# Create table for entries text conflict that need to be resolved
egg_text_conflict_data_v1 <- All_data_v2 %>% # egg_text_conflict: 39 on 2019/09/24
  left_join(egg_status_v1, by = c("Date", "location", "Eggs" = "Eggs_status")) %>%
  filter(egg_text_conflict) %>%
  select(Date, Dorm, Plant, Leaf, Eggs)
#write_csv(egg_text_conflict_data, "egg_text_conflict_data.csv")

# Create table for entries with missing data that need to be resolved
egg_status_missing_data_v1 <- All_data_v2 %>% # egg_status_missing: 219 on 2019/09/24
  left_join(egg_status_v1, by = c("Date", "location", "Eggs" = "Eggs_status")) %>%
  filter(egg_status_missing, !grepl('NA|M', Female_ID)) %>%
  select(Date, Dorm, Plant, Leaf, Eggs)
#write_csv(egg_status_missing_data, "egg_status_missing_data.csv")

@@

#write_csv(egg_status_missing_data_eggs_revised, "egg_status_missing_data_eggs_revised_20200326.csv")
```

@@

#write_csv(egg_status_missing_data_eggs_revised, "egg_status_missing_data_eggs_revised_20200326.csv")

@@

nymph_status_to_resolve <- nymph_status  %>% # should only include a handful of entries with "?" ambiguous entries for N1 or N2 (because did the revions by hand on 2019/09/26)
  group_by(row_number()) %>%
    filter(any(N1_need_filling, N1_text_conflict, N1_status_missing,
               N2_need_filling, N2_text_conflict, N2_status_missing,
               N5_need_filling, N5_text_conflict, N5_status_missing)) %>%
  ungroup()
# write_csv(nymph_status_to_resolve, "nymph_status_to_resolve.csv") ## manually added this to "nymph_status revisions" sheet in "Watershed_field_data_2019_ALL_local" and manually revised raw data 2019/09/26

@@

# checking number of ambiguous entries that need to be resolved
nymph_text_conflict_data <- All_data_v4 %>% # nymph_text_conflict: 10 on 2019/09/24; 0 on 2019/09/26
  left_join(N1_status, by = c("Date", "location", "N1")) %>%
  left_join(N2_status, by = c("Date", "location", "N2")) %>%
  left_join(N5_status, by = c("Date", "location", "N5")) %>%
  group_by(row_number()) %>% 
    filter(any(c(N1_text_conflict, N2_text_conflict, N5_text_conflict))) %>%
  ungroup() %>%
  select(Date, Dorm, Plant, Leaf, Female_ID, Eggs, N1, N2, N5, nymph_origin, Notes, N1_text_conflict, N2_text_conflict, N5_text_conflict)
#write_csv(egg_text_conflict_data, "egg_text_conflict_data.csv")

nymph_status_missing_data <- All_data_v4 %>% # nymph_status_missing: 17 on 2019/09/24; 5 on 2019/09/26 (all ok to ignore)
  left_join(N1_status, by = c("Date", "location", "N1")) %>%
  left_join(N2_status, by = c("Date", "location", "N2")) %>%
  left_join(N5_status, by = c("Date", "location", "N5")) %>%
  group_by(row_number()) %>% 
  filter(any(c(N1_status_missing, N2_status_missing, N5_status_missing))) %>%
  ungroup() %>%
  select(Date, Dorm, Plant, Leaf, Female_ID, Eggs, N1, N2, N5, nymph_origin, Notes, N1_status_missing, N2_status_missing, N5_status_missing)
#write_csv(egg_status_missing_data, "egg_status_missing_data.csv")

@@

## 2.9 - Manually check hatch dates using the "not hatched" notes from field using `check_hatch_date` sheet
Use the "not hatched" notes (from original field notes; typically b/c nymphs present but they are immigrants, not hatched from the focal clutch) to inform first pass of hatch date revisions
Again these changes were specified in the `check_hatch_date` sheet and are applied here without changing the original data file.
```{r}
# Create `check_hatch_date` sheet for checking "not hatched" notes and suggesting manual revisions
check_hatch_date <- All_data_v4  %>%
  filter(grepl("not", Notes, ignore.case = TRUE), grepl("hatch", Notes, ignore.case = TRUE)) %>%
  distinct(location)
#write_csv(check_hatch_date, "check_hatch_date_tmp.csv")
```

@@

#write_csv(ant_text_conflict, "ant_text_conflict.csv") # resolved 2019-10-08

@@

#write_csv(short_unhatched_days_check, "short_unhatched_days_check_tmp.csv")

@@

#write_csv(never_hatched_check, "never_hatched_check_tmp.csv")

#write_csv(clutch_status_stats_v1, "egg_measures_check_tmp.csv")
```

@@

# 6 - Saving data for downstream analysis
